---
title: "Simulation_p_value_linear regression"
author: "Nguyen Truong"
date: "2025-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction
This exercise and solution in the course *Linear model* of the *Master in Statistics and Data Science* at University of Hasselt. I re-write it as practicing. 
**Set up a simulation study to demonstrate that the following hypothesis testing procedure does not control the type I error rate at the nominal $\alpha$ level**. 
*Explaining the question* 
- *The following hypothesis testing procedure* means: if $\hat{beta1}$ < 0, set the alternative hypothesis to H1:$\hat{beta1}$ < 0, otherwise set the alternative hypothesis to H1:$\hat{beta1}$ > 0. 
- *The following hypothesis testing procedure does not control the type I error rate at the nominal $\alpha$ level* means: when we repeat the simulation many times (1.000), the rate/proportion of the type 1 error <0.05 is more often than 0.05    

*Solution* 
- As the requirement in the question, we simulate a data under the null hypothesis (H0:$\hat{beta1}$ = 0) and then compute the p_value for 2 cases of one-side alternative hypothesis above. This p_value is the type I error because type I error is probability to reject H0 when H0 is true.  
- We repeat the simulation many times (1.000) and then compute the proportion/rate of p_value < 0.05. If this rate > 0.05, we say the the hypothesis testing procedure dose not control the type I error rate at 0.05 

```{r}
rm(list = ls())

set.seed(12345)
N <- 10000   # number of repeat experiment
x <- c(165, 170, 175, 180, 185) #create x values (5 subjects) from the father's heights in Galton's experiment
#X <- cbind(1,x) #design the matrix of x
pvalue <- c()

for (i in 1:N) {
  y <- 90 + 0*x + rnorm(n = 5,mean = 0, sd = 5) #random sample of 5 outcomes
  dat <- data.frame(x,y)
  m <- lm(y ~ x, data = dat)
  beta1Hat <- coef(m)[2]
  if(beta1Hat<0) {
    p<- summary(m)$coef[2,4]/2   # if H1: beta1Hat<0, so t<0, the p_value needs to be divided by 2
  }
  if(beta1Hat>0) {
    p <- summary(m)$coef[2,4]/2  # if H1: beta1Hat>0, so t>0, the p_value needs to be divided by 2
  }
  pvalue <- c(pvalue,p) 
}

#compute the rate of the empirical type I error for alpha = 0.05
mean(pvalue < 0.05)

#histogram
hist(pvalue)
hist(pvalue, cex.lab = 1.5, cex.axis = 1.5, xlim = c(0,1))


# empirical type I error rate for alpha=0.
mean(pvalues<0.05)


# histogram 
hist(pvalues, cex.lab=1.5, cex.axis=1.5,
     xlim=c(0,1))

```
note: This simulation study shows that the type I error rate is approximated by 0.0982, which is larger than the nominal value of alpha =0.05
 
